---
title: "Introduction to BFRdist"
author: "Daniel Luo and Jennifer Hanna"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BFRdist}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `BFRdist` package provides computational tools and visualizations for the BFR distribution, a novel, flexible distribution defined on the unit interval (0, 1). It is primarily designed for the analysis of the Generalized Radar Vegetation Index (GRVI) from SAR data.

This vignette demonstrates how to use the core distribution functions, fit empirical data, and visualize the results using the package's object-oriented S3 methods.

## 1. Setup and Core Functions

First, load the package:

```{r load-lib}
library(BFRdist)
```

The package implements the standard suite of `d/p/q/r` functions for the BFR distribution. You can calculate the density (`dbfr`), cumulative probability (`pbfr`), and quantiles (`qbfr`), or generate random deviates (`rbfr`).

```{r core-functions}
# Set shape parameters
alpha_val <- 1.5
beta_val <- 2.0

# Calculate the density at x = 0.5
dbfr(0.5, alpha = alpha_val, beta = beta_val)

# Calculate the cumulative probability up to q = 0.5
pbfr(0.5, alpha = alpha_val, beta = beta_val)

# Find the 50th percentile (median)
qbfr(0.5, alpha = alpha_val, beta = beta_val)
```

## 2. Generating Data and Calculating Summary Statistics

Because the BFR distribution operates on the unit interval, all observations must strictly be between 0 and 1. We can simulate a synthetic GRVI dataset using `rbfr()` and calculate its Galton Skewness and Moors Kurtosis.

```{r simulate-data}
set.seed(123)

# Generate 500 random observations
synthetic_grvi <- rbfr(n = 500, alpha = 2.0, beta = 3.5)

# Calculate summary statistics
galton_skewness(synthetic_grvi)
moors_kurtosis(synthetic_grvi)
```

## 3. Fitting the BFR Distribution

To estimate the alpha and beta parameters from your data, use the `fit_bfr()` function. This function uses a fast, compiled C++ backend to generate initial quantile-based estimates before applying Maximum Likelihood Estimation (MLE) or the Method of Moments (MME). 

The function returns an object of class `bfr`.

```{r fitting}
# Fit the data using Maximum Likelihood
bfr_model <- fit_bfr(synthetic_grvi, method = "mle")
```

## 4. Exploring the Fitted Model (S3 Methods)

The `BFRdist` package includes custom S3 methods to easily extract metrics, simulate new data, and visualize your fitted `bfr` object.

### Print and Summary
The `print()` method outputs the basic parameter estimates, while `summary()` provides standard errors and goodness-of-fit metrics (Log-Likelihood, AIC, and BIC).

```{r summary-methods}
print(bfr_model)

summary(bfr_model)
```

### Predict
You can simulate new observations based on your fitted model using the `predict()` method.

```{r predict-method}
# Simulate 5 new predictions
predict(bfr_model, n.ahead = 5)
```

### Plotting
The `plot()` method generates a Pearson/Cullen Frey-like diagram using `ggplot2`, leveraging the Galton skewness and Moors kurtosis to evaluate the fit.

```{r plot-method, fig.width=7, fig.height=5}
# Generate the diagnostic plot
plot(bfr_model)
```